Seasonality in dominant upper-ocean dynamics at submesoscales

### 2016-05-09 writeup/rocha_etal.tex
Improves introduction and adds appendix

\keypoints{\item Submesoscale (10-100 km) vertical vorticity and lateral strain peaks in late
                 Winter/early Spring;
                 divergence and stratification peaks in late Summer/early Fall.
           \item Quasi-balanced eddies dominate the
                  variability in late Winter/early Spring; inertia-gravity waves
                  dominate in late Summer/early Fall.
            \item These results have implications for the diagnosis of
                  surface velocity from high-resolution altimeters.
                  }

@@

\subsection{Motivation}
Recent studies have suggested that submesoscale upper-ocean flows
undergo strong seasonal cycle \citep{sasaki_etal2014,qiu_etal2014,
callies_etal2015, thompson_etal2016,buckingham_etal2016}.

Observations at submesoscales are limited.

SWOT and COMPIRA will provide submesoscale-resolving SSH measurements
with global coverage.

\subsection{This paper}
Using the output of a $1/24^\circ$ state estimate, we show that (vertical)
vorticity and (horizontal) divergence undergo strong seasonal cycle that are
out of phase. This seasonal cycle is associated with changes in the upper-ocean
stratification. As previously argued, in late Winter/early Spring,
deep mixed-layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation at mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
The divergence part of the flow, however, peaks in late Summer/early Fall,
when the upper-ocean is strongly stratified. Hence these results suggest a
strong seasonal modulation of the dominant upper-ocean submesoscale dynamics
with the dominance quasi-balanced turbulence in late Winter/early Spring and
quasi-linear inertia-gravity waves in late Summer/early Fall.

@@

%  this should go in a supplementary material
%\appendix
%\section*{How well does the model captures high-frequency modes?}
% \begin{figure*}[ht]
%   \begin{center}
%     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
%  \caption{A comp.}
%  \label{fig4}
%  \end{center}
%\end{figure*}

>>>>

\keypoints{\item Submesoscale (10-100 km) vertical vorticity and horizontal divergence
                 divergence undergo a strong seasonal cycle that are out-of-phase.
           \item Quasi-balanced  turbulence dominates the
                  variability in late Winter/early Spring; inertia-gravity waves
                  dominate in late Summer/early Fall.
            \item These results have implications for the diagnosis of
                  surface velocity from high-resolution altimeters.
                  }

@@

Recent interest in upper-ocean dynamics has focused on the strong seasonal
cycle of shallow baroclinic instabilities and their role in submesoscale (1-100 km)
turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
have also suggested that inertia-gravity waves contribute significantly
to the near-surface variability at scale submesoscales \citep{richman_etal2012,
buhler_etal2014,rocha_etal2016}.

Using the output of a $1/24^\circ$ global simulation, we show that the near-surface
expression of inertia-gravity waves undergo strong seasonal cycle that is out-of-phase
with the seasonal cycle of the balanced variability. This seasonal cycle is associated
with changes in the upper-ocean
stratification. Consistent with previous studies,
deep late Winter mixed-layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
The horizontally divergent part of the flow, however, peaks in late Summer/early Fall,
when the upper-ocean is strongly stratified. Hence these results suggest a
strong seasonal modulation of the dominant upper-ocean submesoscale dynamics:
quasi-balanced turbulence in late Winter/early Spring and
quasi-linear inertia-gravity waves in late Summer/early Fall.

@@

%  this should go in a supplementary material
\appendix
\section{How well does the model captures high-frequency modes?}
 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
  \caption{A comparison.}
  \label{figA1}
  \end{center}
\end{figure*}
